Multi-fluids for Representing Sub-filter-scale Convection

### 2020-03-27 submit2/manuscript.tex
Responded to more of reviewer 1's comments

\begin{table}[ht]
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)\tabularnewline
$P_{i}$ & Non-hydrostatic pressure of fluid $i$ $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$)\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$\tabularnewline
$b$ & Buoyancy defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$\gamma$ & Coefficient for setting the pressure local for each fluid ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
\end{tabular}
\caption{Definitions of variables for the multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations; convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore for simplicity we focus on the
\remove[HW]{multi-fluid} Boussinesq equations without background stratification:
\note[HW]{This is additional material until the start of 2.1}
\begin{eqnarray}
\frac{D\mathbf{u}}{Dt}+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{Db}{Dt} & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}. 
These equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere and satisfy:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
Applying the conditional average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
As we are considering the Boussinesq equations we assume that variations in density only affect the buoyancy terms so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

@@

Next we will consider conditionally averaging the terms of the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \mathbf{u} \cdot \nabla \mathbf{u}} &=& \sigma_i \mathbf{u}_i \cdot \nabla \mathbf{u}_i + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \mathbf{u} \cdot \nabla b} &=& \sigma_i \mathbf{u}_i \cdot \nabla b_i + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.

Since viscosity ($\nu$) and thermal diffusivity ($\alpha$) are constant, conditional averaging of the diffusion terms is straightforward:
\begin{eqnarray}
\overline{I_i \nabla \cdot \nu \nabla \mathbf{u}} &=& \sigma_i \nabla\cdot\bigl(\nu \nabla\mathbf{u}_i \bigr) \\
\overline{I_i \nabla \cdot \alpha \nabla b} &=& \sigma_i \nabla\cdot\bigl(\alpha \nabla b_i \bigr).
\end{eqnarray}

Unlike \citeA{TWV+18} we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:
\begin{equation}
\sigma_i \nabla P_i = \overline{I_i \nabla P}.
\end{equation}
Combining these gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{D_{i}\mathbf{u}_{i}}{Dt}+\nabla P_{i}
& = &
b_{i}\mathbf{k}+\nu\nabla^{2}\mathbf{u}_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(\mathbf{u}_{ji}^{T}-\mathbf{u}_{i}\right)-S_{ij}(\mathbf{u}_{ij}^{T}-\mathbf{u}_{i})-\mathbf{D}_{ij}\right\}
\label{eq:mom}\\
\frac{D_{i}b_{i}}{Dt}
& = &
\alpha\nabla^{2}b_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(b_{ji}^{T}-b_{i}\right)-S_{ij}\left(b_{ij}^{T}-b_{i}\right)\right\} \label{eq:b}
\end{eqnarray}
where $D_{i}\big/Dt=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$ is the total derivative with respect to fluid $i$, $\mathbf{u}_{ij}^T$ is the velocity of the fluid transferred from $i$ to $j$, $b_{ij}^T$ is the buoyancy of the fluid transferred from $i$ to $j$ and $\mathbf{D}_{ij}$ is the drag of fluid $j$ on fluid $i$. It is not clear that drag between fluids is appropriate for this form of the equations but it is commonly used in multi-phase modelling \cite<eg>{?} so we will investigate sensitivity to drag in section \ref{sec:results}.

@@

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.

We will assume that pressure in each fluid is given by a parameterised
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating fluidised beds of granular material
\cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}.
In these methods, advection equations are written for the volume fraction,
$\sigma_{i}$, and conservation equations are written for mass fraction,
$\sigma_{i}\rho_{i}$. In our notation, ignoring transfers between
fluids, the \citeA{BN86} equations would be:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)\label{eq:BNcompaction}\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}-\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot\mathbf{u}_{i}\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a free coefficient. Dimensional analysis in section
\ref{subsec:dimAnal} will inform the choice of $\gamma$ and sensitivity
to $\gamma$ will be evaluated in section \ref{sec:results}.

@@

We consider dry air without turbulence so use the dynamic entrainment
described by \citeA{HC51},\citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise}
\end{cases}
\end{equation}
which is necessary to stop air in one fluid from accumulating when
vertical motion ceases (at the top of a rising plume or at the bottom
of descending air). We will also test a common form of lateral entrainment
\cite{DBF+13}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}\implies S_{01}=w_{1}\frac{0.2}{r_{c}}
\end{equation}
where $\varepsilon$ is the well know fractional entrainment rate
and $r_{c}$ is the cloud or plume radius.

@@

The different fluids may not be well mixed so the fluid transferred
may not have the mean properties of the fluid it is leaving \cite<as was assumed by>{WM19}.
In fact, the most buoyant air should be transferred from fluid 0 to
fluid 1 and vice versa, and the air with least downward momentum should
be transferred from 0 to 1 and vice versa. The properties of the fluids
transferred should depend on the modelling of sub-grid-scale variability
which is beyond the scope of this paper. In section \ref{sec:results}
we conditionally average high resolution solutions based on the sign
of $w$ so that $w<0$ air is in fluid 0 and $w>0$ air in fluid 1.
Therefore the air that is transferred at the interface will have $w=0$.
For two dimensional multi-fluid simulations we assume that the horizontal
velocity transferred is equal to the mean horizontal velocity of the
fluid transferred from. Therefore
\begin{equation}
\mathbf{u}_{ij}^{T}=\begin{pmatrix}u_{i}\\
v_{i}\\
0
\end{pmatrix}.
\end{equation}
For buoyancy, following \citeA{TEB19}, we use:
\begin{equation}
b_{ij}^{T}=\theta_{b}b_{i}+(1-\theta_{b})b_{j}
\end{equation}
and present results using $\theta_{b}=\frac{1}{2}$ and $\theta_{b}=1$.
Note that $\theta_{b}\ne1$ implies that the buoyancy of the fluid
transferred depends on the properties of the receiving fluid which
is not logical and can lead to unbounded values of $b_{i}$ in the
fluid that is loosing mass (as demonstrated in section \ref{sec:results}).

\subsection{Dimensional Analysis\label{subsec:dimAnal}}

@@

Next a Poisson equation is solved to calculate the total pressure,
$P$ to ensure that the total velocity is divergence free. The pressure
equation also provides updates for the velocity flux, $u_{i}$. First
an intermediate velocity flux is calculated with partial updates and
updates from a previous value of $p_{i}$:
\begin{eqnarray}
u_{i}^{\prime}= & u_{i}^{n}+\frac{\Delta t}{2} & \biggl(\left(-\nabla\cdot(\mathbf{u}_{i}\mathbf{u}_{i})+\mathbf{u}_{i}\nabla\cdot\mathbf{u}_{i}+b_{i}\mathbf{k}\right)_{f}^{\ell}\cdot\mathbf{S}-\nabla_{f}p_{i}^{\ell}\\
 &  & +\left(-\nabla\cdot(\mathbf{u}_{i}\mathbf{u}_{i})+\mathbf{u}_{i}\nabla\cdot\mathbf{u}_{i}+b_{i}\mathbf{k}\right)_{f}^{n}\cdot\mathbf{S}-\nabla_{f}(P+p_{i})^{n}\biggr)
\end{eqnarray}
where operator $()_{f}$ means linear interpolation from cell centres
onto cell faces and $\nabla_{f}P$ is a compact discretisation of
$(\nabla P)_{f}\cdot\mathbf{S}$: ie the normal component of the pressure
gradient calculated from just the values of pressure either side of
the face. Once we have calculated $P^{n+1}$, the velocity flux (without
momentum transfers) is updated from the back-substitution:
\begin{equation}
u_{i}^{\prime\prime}=u_{i}^{\prime}-\frac{\Delta t}{2}\nabla_{f}P^{n+1}.\label{eq:backSub}
\end{equation}
The Poisson equation for $P^{n+1}$ is found by multiplying each eqn
(\ref{eq:backSub}) by $\sigma_{i}$, summing over all fluids and
taking the divergence. Knowing that $\nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0$,
$\sum_{i}\sigma_{i}=1$ and $\sum_{i}\sigma_{i}p_{i}=0$ gives:
\begin{equation}
\nabla\cdot\sum_{i}\sigma_{i}u_{i}^{\prime}-\frac{\Delta t}{2}\nabla^{2}P^{n+1}=0
\end{equation}
which is solved to find $P^{n+1}$.

>>>>

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations; convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore for simplicity we focus on the
\remove[HW]{multi-fluid} Boussinesq equations without background stratification:
\note[HW]{This is additional material until the start of 2.1}
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}.

\begin{table}
\begin{tabular}{c>{\raggedright}p{0.85\textwidth}}
$\sigma_{i}$ & Volume fraction of fluid $i$ so that $\sum_{i}\sigma_{i}=1$\tabularnewline
$\mathbf{u}$ & 
Velocity of the single-fluid equations  ($\text{m}\text{s}^{-1}$) and velocity averaged over all fluids of the multi-fluid equations
\tabularnewline
$\mathbf{u}_{i}$ & Velocity of fluid $i$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$w_{i}$ & Vertical component of $\mathbf{u}_{i}$ ($\text{m}\text{s}^{-1}$)\tabularnewline
$S_{ij}$ & Transfer from fluid $i$ to $j$ ($\text{s}^{-1}$)
\tabularnewline
$P$ & Non-hydrostatic pressure of the single-fluid equations $=p^{\prime}/\rho_{r}$ ($\text{m}^{2}\text{s}^{-2}$) and pressure averaged over all fluids of the multi-fluid equations
\tabularnewline
$P_{i}$ & Non-hydrostatic pressure of fluid $i$ 
\tabularnewline
$p_{i}$ & $P_i - P$
\tabularnewline
$\psi_{ij}^{T}$ & Value of variable $\psi$ transferred from fluid $i$ to $j$\tabularnewline
$b$ & Buoyancy of single-fluid equations defined as -$g\rho^{\prime}/\text{\ensuremath{\rho}}_{r}$
where $\rho^{\prime}$ are departures in density from a horizontally
uniform reference, $\rho_{r}$ ($\text{m}\text{s}^{-2}$) and buoyancy averaged over all fluids of the multi-fluid equations
\tabularnewline
$b_{i}$ & Buoyancy of fluid $i$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$D_{i}\big/Dt$ & Total derivative with respect to fluid $i$ $=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$\tabularnewline
$\mathbf{D}_{ij}$ & Drag on fluid $i$ from fluid $j$ ($\text{m}\text{s}^{-2}$)\tabularnewline
$\gamma$ & Coefficient for setting the pressure local for each fluid ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$C_{D}$ & Drag coefficient $=0$ or 1\tabularnewline
$r_{c}$ & Plume radius used for defining the drag between fluids (m)\tabularnewline
$\alpha$ & Diffusivity of buoyancy ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
$\nu$ & Viscosity ($\text{m}^{2}\text{s}^{-1}$)\tabularnewline
\end{tabular}
\caption{Definitions of variables for the single- and multi-fluid Boussinesq equations.\label{tab:defns}}
\end{table}

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere and satisfy:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
Applying the conditional average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
As we are considering the Boussinesq equations we assume that variations in density only affect the buoyancy terms so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

@@

Next we will consider conditionally averaging the terms of the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.

\note[HW]{Dan, can we do better than this?}
Lots of terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. Therefore we will use simply:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray}

Unlike \citeA{TWV+18} we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i.
\end{equation}

Combining these gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

Eqns ({eq:momFlux}) and ({eq:bFlux}) are solved in advective form in which the mass transfer terms are more complicated:
\begin{eqnarray}
\frac{D_{i}\mathbf{u}_{i}}{Dt}+\nabla P_{i}
& = &
b_{i}\mathbf{k}+\nu\nabla^{2}\mathbf{u}_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(\mathbf{u}_{ji}^{T}-\mathbf{u}_{i}\right)-S_{ij}(\mathbf{u}_{ij}^{T}-\mathbf{u}_{i})-\mathbf{D}_{ij}\right\}
\label{eq:mom}\\
\frac{D_{i}b_{i}}{Dt}
& = &
\alpha\nabla^{2}b_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(b_{ji}^{T}-b_{i}\right)-S_{ij}\left(b_{ij}^{T}-b_{i}\right)\right\} \label{eq:b}
\end{eqnarray}
where $D_{i}\big/Dt=\partial/\partial t+\mathbf{u}_{i}\cdot\nabla$ is the total derivative with respect to fluid $i$,

@@

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. \change[HW]{It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.}{For single-fluid equations, divergence would lead to pressure anomalies which would act as a negative feedback on the divergence. However for multi-fluids if one of the volume fractions is too small then it cannot force pressure anomalies big enough to stabilise the divergence. Use of separate pressure for each fluid is proposed to ensure stability.}

We will assume that pressure in each fluid is given by a parameterised
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model.}

\subsubsection{\add[HW]{The \protect\citeA{BN86} model}}

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a method suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

\subsubsection{\add[HW]{Proposed Model for Pressure Differences between Fluids}}

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot(\sigma_i \mathbf{u}_{i})
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. 
This parameterisation of $p_i$ has $\sigma_i$ inside the divergence in order to ensure that $\sum_i p_i =0$ and $\therefore$ that $\sum_i P_i = P$.}

\change[HW]{Dimensional}{Scale} analysis in section
\ref{subsec:dimAnal} will inform the choice of $\gamma$ and sensitivity
to $\gamma$ will be evaluated in section \ref{sec:results}.

@@

We consider dry air without turbulence so use the dynamic entrainment
described by \citeA{HC51},\citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise}
\end{cases}
\end{equation}
which is necessary to stop air in one fluid from accumulating when
vertical motion ceases (at the top of a rising plume or at the bottom
of descending air). We will also test a common form of lateral entrainment
\cite{DBF+13}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}
\end{equation}
where $\varepsilon$ is the well know fractional entrainment rate
and $r_{c}$ is the cloud or plume radius. Fractional entrainment is related to mass transfers by:
\begin{equation}
S_{01} = w_{1}\varepsilon.
\end{equation}

@@

The different fluids may not be well mixed so the fluid transferred
may not have the mean properties of the fluid it is leaving \cite<as was assumed by>{WM19}.
In fact, the most buoyant air should be transferred from fluid 0 to
fluid 1 and vice versa, and the air with least downward momentum should
be transferred from 0 to 1 and vice versa. The properties of the fluids
transferred should depend on the modelling of sub-grid-scale variability
which is beyond the scope of this paper. In section \ref{sec:results}
we conditionally average high resolution solutions based on the sign
of $w$ so that $w<0$ air is in fluid 0 and $w>0$ air in fluid 1.
Therefore the air that is transferred at the interface will have $w=0$.
For two dimensional multi-fluid simulations we assume that the horizontal
velocity transferred is equal to the mean horizontal velocity of the
fluid transferred from. Therefore
\begin{equation}
\mathbf{u}_{ij}^{T}=\begin{pmatrix}u_{i}\\
v_{i}\\
0
\end{pmatrix}.
\end{equation}
For buoyancy, following \citeA{TEB19}, we use:
\begin{equation}
b_{ij}^{T}=\theta_{b}b_{i}+(1-\theta_{b})b_{j}
\end{equation}
and present results using $\theta_{b}=\frac{1}{2}$ and $\theta_{b}=1$.
\add[HW]{
$\theta_{b}=1$ is described as an upstream approximation by \protect\citeA{Yano14} and is widely used for entrainment \protect\cite<eg.>{AK67,AS74}.
}
Note that $\theta_{b}\ne1$ implies that the buoyancy of the fluid
transferred depends on the properties of the receiving fluid which
is not logical and can lead to unbounded values of $b_{i}$ in the
fluid that is loosing mass (as demonstrated in section \ref{sec:results}).

\subsection{\change[HW]{Dimensional}{Scale} Analysis\label{subsec:dimAnal}}

@@

Next a Poisson equation is solved to calculate the total pressure,
$P$ to ensure that the total velocity is divergence free. \add[HW]{This is the standard approach for ensuring divergence constraints \protect\cite<eg>{SKW14}.} The pressure
equation also provides updates for the velocity flux, $u_{i}$. First
an intermediate velocity flux is calculated with partial updates and
updates from a previous value of $p_{i}$:
\begin{eqnarray}
u_{i}^{\prime}= & u_{i}^{n}+\frac{\Delta t}{2} & \biggl(\left(-\nabla\cdot(\mathbf{u}_{i}\mathbf{u}_{i})+\mathbf{u}_{i}\nabla\cdot\mathbf{u}_{i}+b_{i}\mathbf{k}\right)_{f}^{\ell}\cdot\mathbf{S}-\nabla_{f}p_{i}^{\ell}\\
 &  & +\left(-\nabla\cdot(\mathbf{u}_{i}\mathbf{u}_{i})+\mathbf{u}_{i}\nabla\cdot\mathbf{u}_{i}+b_{i}\mathbf{k}\right)_{f}^{n}\cdot\mathbf{S}-\nabla_{f}(P+p_{i})^{n}\biggr)
\end{eqnarray}
where operator $()_{f}$ means linear interpolation from cell centres
onto cell faces and $\nabla_{f}P$ is a compact discretisation of
$(\nabla P)_{f}\cdot\mathbf{S}$: ie the normal component of the pressure
gradient calculated from just the values of pressure either side of
the face. Once we have calculated $P^{n+1}$, the velocity flux (without
momentum transfers) is updated from the back-substitution:
\begin{equation}
u_{i}^{\prime\prime}=u_{i}^{\prime}-\frac{\Delta t}{2}\nabla_{f}P^{n+1}.\label{eq:backSub}
\end{equation}
The Poisson equation for $P^{n+1}$ is found by multiplying each eqn
(\ref{eq:backSub}) by $\sigma_{i}$, summing over all fluids and
taking the divergence. Knowing that $\nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0$,
$\sum_{i}\sigma_{i}=1$ and $\sum_{i}\sigma_{i}p_{i}=0$ gives:
\begin{equation}
\nabla\cdot\sum_{i}\sigma_{i}u_{i}^{\prime}-\frac{\Delta t}{2}\nabla^{2}P^{n+1}=0
\end{equation}
which is solved to find $P^{n+1}$.
